---
title: "Exercise 11.2"
author: "Andrew Jordan"
date: '2022-06-02'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
knitr::opts_knit$set(root.dir = 'C:/Users/Andrew/StatClass/dsc520/')
knitr::opts_chunk$set(echo = TRUE, warning= FALSE, message=FALSE)
getwd()
```
## Question 1: Introduction to Machine Learning
```{r}
binary_df <- read.csv("data/binary-classifier-data.csv")
trinary_df <- read.csv("data/trinary-classifier-data.csv")
head(binary_df)
head(trinary_df)
```
```{r}
library(ggplot2)
ggplot(binary_df, aes(x = x, y = y, color = label)) + geom_point() + ggtitle("Binary Classifier Data")
```
```{r}
ggplot(trinary_df, aes(x = x, y = y, color = label)) + geom_point() + ggtitle("Trinary Classifier Data")
```

```{r}
binary_dm <- as.matrix(binary_df)
binary_k3 <- kmeans(binary_dm, 3)
binary_k5 <- kmeans(binary_dm, 5)
binary_k10 <- kmeans(binary_dm, 10)
binary_k15 <- kmeans(binary_dm, 15)
binary_k20 <- kmeans(binary_dm, 20)
binary_k25 <- kmeans(binary_dm, 25)
```


```{r}
trinary_dm <- as.matrix(trinary_df)
trinary_k3 <- kmeans(trinary_dm, 3)
trinary_k5 <- kmeans(trinary_dm, 5)
trinary_k10 <- kmeans(trinary_dm, 10)
trinary_k15 <- kmeans(trinary_dm, 15)
trinary_k20 <- kmeans(trinary_dm, 20)
trinary_k25 <- kmeans(trinary_dm, 25)
```

At this point professor, I am unsure of how to compute the accuracy of the resulting models for each value of k. I've gone through the textbook and additional materials, but I believe I am either overlooking or not understanding something essential to my comprehending checking the KNN models' accuracy. 

## Question 2: Clustering
```{r}
cluster_df <- read.csv("data/clustering-data.csv")
head(cluster_df)
ggplot(cluster_df, aes(x = x, y = y, color = "red")) + geom_point() + ggtitle("Clustering Data")
```
```{r}
cluster_dm <- as.matrix(cluster_df)
cluster_k2 <- kmeans(cluster_dm, 2)
cluster_k3 <- kmeans(cluster_dm, 3)
cluster_k4 <- kmeans(cluster_dm, 4)
cluster_k5 <- kmeans(cluster_dm, 5)
cluster_k6 <- kmeans(cluster_dm, 6)
cluster_k7 <- kmeans(cluster_dm, 7)
cluster_k8 <- kmeans(cluster_dm, 8)
cluster_k9 <- kmeans(cluster_dm, 9)
cluster_k10 <- kmeans(cluster_dm, 10)
cluster_k11 <- kmeans(cluster_dm, 11)
cluster_k12 <- kmeans(cluster_dm, 12)
```
```{r}
library(factoextra)

fviz_cluster(cluster_k2, cluster_df, geom = "point")
fviz_cluster(cluster_k3, cluster_df, geom = "point")
fviz_cluster(cluster_k4, cluster_df, geom = "point")
fviz_cluster(cluster_k5, cluster_df, geom = "point")
fviz_cluster(cluster_k6, cluster_df, geom = "point")
fviz_cluster(cluster_k7, cluster_df, geom = "point")
fviz_cluster(cluster_k8, cluster_df, geom = "point")
fviz_cluster(cluster_k9, cluster_df, geom = "point")
fviz_cluster(cluster_k10, cluster_df, geom = "point")
fviz_cluster(cluster_k11, cluster_df, geom = "point")
fviz_cluster(cluster_k12, cluster_df, geom = "point")
```
Professor, this again appears to be the point of my limitation in this assignment. After plotting the kmeans cluster results, I am at a loss as to how to proceed through the next step. I am continuing to search the internet for a solution, but at this time I again must be either overlooking or misunderstanding something from the textbook and additional readings.
